---
title: "HW1"
author: "Zeqiu.Yu"
date: "2022-09-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Homework1{#sec:requirements}
1. Read Chapter 1 in textbook (givens and Hoeting)              FINISHED!
2. Install R and Rstudion in your computer                      FINISHED!
3. Go through the Rmarkdown and R codes in Intro.Rmd            FINISHED!
4. If you take cards numbered from 1-10 and shuffle them, and lay them down in order, what is the probability that at least one card matches its position. For example card 3 comes down third? Using simulation to find the answer and compare it to the theoretical value.                 FINISHED!
\newpage

## Q4{#sec:question4}
1. Define a function to judge whether there is at least one card matches its position. Yes = 1, No = 0
```{r ComparingFunction, echo = TRUE, eval = TRUE, warning = FALSE, result = TRUE,include = TRUE}
IdenticalVector <- function(x,y){
  numTrue <-  0
  for(i in 1:length(x)){
    if(x[i]==y[i]){
      numTrue <-  numTrue + 1
    }
  }
  if(numTrue >0){
    return(1)
  }else{
    return(0)
  }
}

```
For n trials, find successful times:
```{r Trials, echo = TRUE, eval = TRUE, warning = FALSE, result = TRUE,include = TRUE}
cPostion <- c(1:10)
for(numTrials in c(10, 100, 1000, 10000,100000)){
numSuccess <-  0
for(i in 1:numTrials){
  afterShuffle <- sample(cPostion, replace = F)
  numSuccess <- numSuccess + IdenticalVector(cPostion, afterShuffle)
}
print(sprintf("After %d trials, the frequency is %.4f", numTrials, numSuccess/numTrials))}

```
And the theoretical value is 
\[
1 - \frac{10!\sum_{i=0}^{n} \frac{(-1)^i}{i!}}{\mathrm{P}_{10}^{10}}
\]
Hence, we can get theoretical result:
```{r TheoreticalTrials, echo = TRUE, eval = TRUE, warning = FALSE, result = TRUE,include = TRUE}
sum = 0
for(i in c(0:10)){
  eachFactor <-  factorial(10)*(-1)^(i)/factorial(i)/factorial(10)
  sum <- sum + eachFactor
}
print(1-sum)

```

