---
title: "HW2"
author: "Zeqiu.Yu"
date: "2022-09-26"
output:
  pdf_document: default
  html_document: default
---


# Question 1

1. Use Simulation to approximate the following integrals. Describe your algorithm first. Attach your code and output.
* \[ \int_{1}^{3}\frac{x}{(1 + x^2)^2} \,dx\ \]
* \[ \int_{-\infty}^{\infty} x^2e^{-x^2} \,dx\ \]
* \[ \int_{-1}^{1}\int_{-1}^{1}|x-y|\;dx\;dy\]
Solutions  

Because we use \[\frac{1}{n}\,\sum_{i = 1}^{n}\,h(x)\] to estimate E[h(X)].We should define h(X) and f(X) for each integral. 
(1) For the first integral
The former intgral can be writtern as
\[ \int_{1}^{3}\frac{1}{2}*2\frac{x}{(1 + x^2)^2} \,dx\ \]
\[X\sim Uniform(1,3),\; f(x) = \frac{1}{2},\; 1<=x<=3\]
\[h(x) = \frac{2x}{(1 + x^2)^2}\]
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
n_list <- c(1000, 10000, 100000,1000000)
for(n in n_list){
  # Sampling
  x <- runif(n,1,3)
  
  s <- 2*x/(1+x^2)^2
  outcome <- 1/n*sum(s)
  print(sprintf("The integral is %.4f with the sample size of %d", outcome, n))
}
```
(2) For the second integral
The former intgral can be writtern as
\[ \int_{0}^{\infty} \frac{1}{2\sqrt{x}}e^{-x} \,dx\ \]
\[X\sim Exp(1),\; f(x) = e^{-x},\; x>0\]
\[h(x) = \frac{1}{2\sqrt{x}}\]
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
n_list <- c(1000, 10000, 100000,1000000)
for(n in n_list){
  # Sampling
  x <- rexp(n,1)
  
  s <- 1/(2*sqrt(x))
  outcome <- 1/n*sum(s)
  print(sprintf("The integral is %.4f with the sample size of %d", outcome, n))
}
```
(3) For the third integral
The former intgral can be writtern as
\[ \int_{-1}^{1}\frac{1}{2}\int_{-1}^{1}\frac{1}{2} 4|x-y|\;dx\;dy\]
\[X\sim Uniform(-1,1),\; Y\sim Uniform(-1,1)\]
\[f(x,y) = \int_{-1}^{1}\int_{-1}^{1}\frac{1}{4}\;dx\;dy,\; x>0\]
\[h(x,y) = 4|x-y|\]
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
n_list <- c(1000, 10000, 100000,1000000)
for(n in n_list){
  # Sampling
  x <- runif(n,-1,1)
  y <- runif(n,-1,1)
  
  s <- 4*abs(x-y)
  outcome <- 1/n*sum(s)
  print(sprintf("The integral is %.4f with the sample size of %d", outcome, n))
}
```

# Question 2
Consider the model given by 
\[X\sim lognormal(0,1)\;\;\;\; and\; logY = 9 + 3logX + e\, , e\sim N(0,1)\]
is independt of X. Use simulation to estimate E[Y/X].  
Let Z be the standard normal distribution, then f(x,z) = f(x)f(z) because of the independence of x and z.
\[h(x,z) = \frac{Y}{X} = e^{9+z}x^2\]
\[E[Y/X] = \iint_{X,Z} h(x,z)f(x,z)\;dxdy =\int_Xx^2f(x)dx\int_Ze^{9+z}f(z)dz\]


```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
n_list <- c(1000, 10000, 100000,1000000,10000000)

for(n in n_list){
  # Sampling
  x <- rlnorm(n,0,1) 
  z <- rnorm(n,0,1)
  s1 <- sum(x^2)/n
  s2 <- sum(exp(9+z))/n
  outcome <- s1*s2
  print(sprintf("The integral is %.4f with the sample size of %d", outcome, n))
}
```

