---
title: "HW5"
author: "Zeqiu.Yu"
date: "2022-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework 5  
## Question 1:  
Set seed of the random number generator to be 12345 using set.seed(12345). Generate 100 random numbers from exponential distribution with mean=1. Let $s^2$ be the sample variance, an estimator of the population variance (6 pts).  
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
set.seed(12345)
SampleSzie1 <- 100
MySample1 <- rexp(SampleSzie1, 1)
```  
1. Use bootstrap method to estimate the variance of $s^2$.  
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
set.seed(12345)
B <-  100000
Samplevariances1 <-  numeric(B)
for(i in 1:B){
  Samplevariance <- sample(MySample1,100,replace=TRUE)
  Samplevariances1[i] <- var(Samplevariance)
}
 print(sprintf("The estimation of the variance of s^2 using bootstrap is %.3f", var(Samplevariances1)))
```  
2. Using Monte-Carlo method to generate 1000 random samples of size 100, compute variance of $s^2$ evaluated based on the 1000 random samples , and compare it with results from part (a).  
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
set.seed(12345)
Sample.num <-  1000
Samplevariances2 <-  numeric(Sample.num)
for(i in 1:Sample.num){
  x <- rexp(100,1)
  Samplevariances2[i] <- var(x)
}
 print(sprintf("The estimation of the variance of s^2 using MC is %.3f.", var(Samplevariances2)))
```  
 It is much smaller than the result in (a), it has higher efficiency.  
 
## Question 2:  
Use the same seed as in 1, generate a random sample with n=15 random observations from exponential distribution with mean=1 (6 pts).  
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
set.seed(12345)
SampleSzie2 <- 15
MySample2 <- rexp(SampleSzie2, 1)
```  
1. Calculate the sample median, which is an estimator of the population median.  
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
SampleMedian <- median(MySample2)
 print(sprintf("The sample median %.3f", SampleMedian))
```  
2. Use bootstrap (nonparametric, with B=1000) methods to estimate the variance of the estimator for the population median.  
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
set.seed(12345)
B <-  1000
SampleMedians <-  numeric(B)
for(i in 1:B){
  SampleMedian <- sample(MySample2,15,replace=TRUE)
  SampleMedians[i] <- median(SampleMedian)
}
 print(sprintf("The variance of the estimator for the  population median is %.3f using bootstrap", var(SampleMedians)))
```  
3. use Monte Carlo method, e.g. generate 1000 samples of size 15 to estimate the true variance of the median estimator. Compare and comment your results.  
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
set.seed(12345)
Sample.num = 1000
SampleMedians2 <-  numeric(Sample.num)
for(i in 1:Sample.num){
  x <- rexp(15,1)
  SampleMedians2[i] <- median(x)
}
 print(sprintf("The variance of the estimator for the  population median is %.3f using MC", var(SampleMedians2)))
```  
Bootstrap works better when the estimator is a location parameter.  

Question 3:  
Patients with advanced terminal cancer of the stomach and breast were treated with ascorbate in an attempt to prolong survival [87]. Table 9.4 shows survival times (days). Work with the data on the log scale. 
(a.) Use the bootstrap t method to construct 95% confidence intervals for the mean survival time of each group.
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
t1 <- log(c(25,42,45,46,51,103,124,146,340,396,412,876,1112))
t2 <- log(c(24,40,719,727,791,1166,1235,1581,1804,3460,3808))

locator1 <- mean(t1)
locator2 <- mean(t2)
sigma1 <- sqrt(var(t1)/length(t1))
sigma2 <- sqrt(var(t2)/length(t2))
B <- 1000000
MST1s <- numeric(B)
MST2s <- numeric(B)
set.seed(12345)
for(i in 1:B){
  MST1 <- sample(t1,length(t1),replace=TRUE)
  MST2 <- sample(t2,length(t2),replace=TRUE)
  M1 <- mean(MST1)
  M2 <- mean(MST2)
  MST1s[i] <- (M1 - locator1)/sqrt(var(MST1)/length(MST1))
  MST2s[i] <- (M2 - locator2)/sqrt(var(MST2)/length(MST2))
}

q1 <- quantile(MST1s, c(0.025,0.975))
q2 <- quantile(MST2s, c(0.025,0.975))
bounds1 <-  exp(c(locator1 - sigma1*q1[[2]], locator1 - sigma1*q1[[1]]))
bounds2 <- exp(c(locator2 - sigma2*q2[[2]], locator2 - sigma2*q2[[1]]))
names(bounds1) <- c("2.5%", "97.5")
names(bounds2) <- c("2.5%", "97.5")
bounds1
bounds2
```  